A Comparative Quality Metric for Fuzzing with Logic State Coverage

### 2024-03-03 paper/main.tex
paper: de-bucketing..

As it is challenging to directly count interesting program behaviors when
fuzzing produces hundreds of them every second, we propose \emph{logic state
coverage} as a feasible proxy metric of interesting behaviors. A logic state is
a set of all satisfied branches during one execution, bucketed by their hit
counts (larger buckets for higher hit counts). Logic state coverage then
\emph{individually counts} all observed distinct logic states.  One logic state
represents \emph{fewer} program behaviors if they are \emph{more interesting}
(i.e., likely relating to unknown abnormal behaviors), meaning that observing
more interesting behaviors is strongly reflected on larger logic state coverage.

@@

\pp{Interestingness in practice}
%
As the interestingness $P(\mathrm{UA} \mid \mathrm{N})$ is the expectation of an
\emph{unobserved} behavior, there is no direct way to measure interestingness. 
%
However, the practice of untargeted fuzzing suggests that
the resulting program behavior is unlikely to be \emph{unknown abnormal} if the
mutation base is highly repetitive, as it commonly incorporates corpus
minimization \cite{} and branch hit count buckets \cite{}.
%
The rationales are: (i) abnormal behaviors might have been already observed with
fewer branch repetitions (i.e., unlikely \emph{abnormal}) and (ii) if so, it may
correspond to pre-observed less-repetition counterparts (i.e., unlikely
\emph{unknown}).

%The prime example is the branch hit count bucketing in AFL \cite{afl}, where
%the bucket size gets larger roughly by the power of 2 (i.e., 1, 2, 3, 4-7,
%8-15, ...).

@@

\pp{Logic state and logic space}
%
A \emph{logic state} $l$ is a set of all satisfied branches during execution,
each of which is paired with the integer logarithm of the corresponding branch
hit count. Formally, given a branch history $\mathcal{T}_b$ of an execution,
%
\begin{equation}
\label{e:ls}
  l := \{(b, n(b)) \mid b \in \mathcal{T}_b\},
\end{equation}

\noindent where $n(b) = \lfloor \log_2{m(b)} \rfloor$ and $m(b)$ is the hit
count of a branch $b$ in $\mathcal{T}_b$.  Notice that, as it gets larger, a
branch hit count $m(b)$ shares the same integer "bucket" $n(b)$ with
exponentially more hit counts (i.e., $n(b)=x \iff m(b) \in \{2^x, ...,
2^{x+1}-1\}$).

@@

As shown in \autoref{f:prog-to-logic}, a logic state (i.e., a partition in the
program behavior space) may contain various program behaviors, suggesting a
single covered logic state may encompass several different behaviors. However,
in \autoref{s:prop}, we argue that a logic state is representative of
interesting program behaviors as it (i) weighs interesting program behaviors
over uninteresting ones and (ii) only contains the program behaviors of the same
normality.

>>>>

As it is challenging to directly count interesting program behaviors when
fuzzing produces hundreds of them every second, we propose \emph{logic state
coverage} as a feasible proxy metric of interesting behaviors. A logic state is
a set of all satisfied branches during one execution, where logic state coverage
is defined as the \emph{individual count} of all observed distinct logic states.
One logic state represents \emph{fewer} behaviors if they are \emph{more
interesting} (i.e., likely relating to unknown abnormal behaviors), meaning that
observing more interesting behaviors is strongly reflected on larger logic state
coverage.

@@

\pp{Interestingness in practice}
%
As the interestingness $P(\mathrm{UA} \mid \mathrm{N})$ is the expectation of an
\emph{unobserved} behavior, there is no direct way to measure interestingness. 
%
However, the practice of untargeted fuzzing suggests that
the resulting program behavior is unlikely to be \emph{unknown abnormal} if the
mutation base is highly repetitive, as it commonly incorporates corpus
minimization \cite{todo} and branch hit count buckets \cite{todo}.
%
The rationales are: (i) abnormal behaviors might have been already observed with
fewer branch repetitions (i.e., unlikely \emph{abnormal}) and (ii) if so, it may
correspond to pre-observed less-repetition counterparts (i.e., unlikely
\emph{unknown}).

@@

\pp{Logic state and logic space}
%
A \emph{logic state} $l$ is a set of all satisfied branches during execution,
which is essentially the order-ignored version of a branch history. Formally,
given a branch history $\mathcal{T}_b$ of an execution,
%
\begin{equation}
\label{e:ls}
  l := \{b \mid b \in \mathcal{T}_b\},
\end{equation}

@@

As shown in \autoref{f:prog-to-logic}, a logic state (i.e., a partition in the
program behavior space) may contain various program behaviors, suggesting a
single covered logic state may encompass several different behaviors. However,
in \autoref{s:prop}, we explain that a logic state is representative of
interesting program behaviors as it (i) reflects the interestingness of
underlying program behaviors and (ii) only contains the program behaviors of the
same normality.
